---
title: "Ward_abigail_HW7"
author: "Abigail Ward"
date: "8/3/2021"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(ggpubr)
library(boot)
library(lawstat)
```

## Question 1


```{r}
dat<-read.table("/Users/abbyward/Downloads/precip.txt",sep="\t",header = TRUE)

# Change all characters in the variable names to lower case.
names(dat)<-str_to_lower(names(dat))
names(dat)

dat<-filter(dat,year!="2021")

# Replace "Tr".
dat<-mutate_all(dat,str_replace,"Tr","0")
# Count all occurrences of "Tr".
sum(str_detect(unlist(dat),"Tr"))

# function to return TRUE if a string vector x contains no entries with an "*".
no_stars<-function(x){
  sum(str_detect(x,"\\*"))==0
}

# Count asterisks in the data.
sum(str_detect(unlist(dat),"\\*"))

# Identify the rows in the data with at least 1 "*".
all.standard<-apply(dat,1,no_stars)
dat.trim<-dat[all.standard,]
# Count asterisks in the trimmed data.
sum(str_detect(unlist(dat.trim),"\\*"))

#set value type
dat.trim<-mutate_all(dat.trim,as.numeric)
dat.trim[,1]<-as.integer(dat.trim[,1])

sapply(dat.trim,class)

sum(is.na(dat))
which(is.na(dat), arr.ind=TRUE)

setdiff(min(dat.trim$year):max(dat.trim$year),dat.trim$year)
```


### 1.a. 
Both the first 15 and last 15 data points appear normal respectively. All 15 data points fit within the permitted normal region. 

```{r}
dat.s<-filter(dat.trim,year%%3==2)

nrow(dat.s)
nrow(dat.s)-14
(nrow(dat.s)-14):nrow(dat.s)

c(1:15,(nrow(dat.s)-14):nrow(dat.s))

dat.sep<-dat.s[c(1:15,(nrow(dat.s)-14):nrow(dat.s)),]
dat.sep$era<-rep(c("first 15","last 15"),
            times=c(15,15))


ggqqplot(dat.sep$year.total,facet.by=c("dat.sep$era"))


```
## 1.b. 


There is no smooth variation in the data over time. There is no consistent upward or downward trend in the first or last data and therefore the year total values appear to be independent.
```{r}
ggplot(data=dat.sep,aes(x=year,y=year.total,color=era))+geom_line()
```


## 1.c.


Since the P-value is 1.3%, there is strong evidence against the null hypothesis that the total annual rainfalls in the early portion and the total annual rainfalls in the recent portion are each i.i.d. samples from Normally distributed populations with equal means, $Normal(\mu,\sigma^2_{early})$ and $Normal(\mu,\sigma^2_{recent})$. This is also re-enforced by parts a and b showing the data is both normal and independent. 

```{r}
t.test(year.total~era,data=dat.sep)

```

## Question 2

The goal in this analysis is to perform the strongest suitable test of whether the precipitation amount differs between October and November.

### 2.a.

The data points all fall within or very close to the permitted normal region in the first plot and there is no trends in the data in the second plot which indicate that the data is both normal and independent. 
With a p-value of abouy 8%, we fail to reject the null hypothesis. 

```{r}
diff<-dat.s$oct-dat.s$nov
ggqqplot(diff)
ggplot(dat.s,aes(x=year,y=oct-nov))+geom_line()

shapiro.test(diff)

```

### 2.b.

With a p value of 0.3064, we fail to reject the null hypothesis that the true difference int the means is equal to 0.


```{r}
x<-dat.s$oct
y<-dat.s$nov
t.test(x,y, paired = TRUE, conf.level = .99)
```


## Question 3
```{r}
load("/Users/abbyward/Downloads/dat_mf.RData")
ggplot(dat.mf, aes(x=HHINCOME, color=as.factor(HHTYPE))) +
  geom_density()
```


### 3.a. 


The household incomes for the male-headed households are not approximately Normally distributed and neither are the household incomes for the female-headed households. The data points for both sets fall outside the permitted normal range. 

```{r}
ggqqplot(dat.mf$HHINCOME[dat.mf$HHTYPE==2])
ggqqplot(dat.mf$HHINCOME[dat.mf$HHTYPE==3])
```


### 3.b. 


Since the p-value is 0.000001427 we reject the null hypothesis that there is no difference in the household incomes for the male-headed households and the household incomes for the female-headed households. This is a test of center as it is comparing the medians of the two data sets.

```{r}
wilcox.test(dat.mf$HHINCOME[dat.mf$HHTYPE==2],
            dat.mf$HHINCOME[dat.mf$HHTYPE==3])
median(dat.mf$HHINCOME[dat.mf$HHTYPE==2])
median(dat.mf$HHINCOME[dat.mf$HHTYPE==3])

```

## 3.c
```{r cache=TRUE}

boot.mean.diff<-function(dat,indices){
  dat.this<-dat[indices,]
  gp2<-dat.this$HHINCOME[dat.this$HHTYPE==2]
  gp3<-dat.this$HHINCOME[dat.this$HHTYPE==3]
  return(c(length(gp2),length(gp3),mean(gp2)-mean(gp3)))
}

# Draw 5000 bootstrap samples stratified by household type. 

samp<-boot(dat.mf,boot.mean.diff,5000,strata=dat.mf$HHTYPE)


unique(samp$t[,1])
unique(samp$t[,2])

# Look at quantiles of the mean difference

quantile(samp$t[,3],c(.025,.975))

# Another interval estimate

boot.ci(samp,type="bca",index=3)



```