---
title: "Ward_Abigail_HW5"
author: "Abigail Ward"
date: "7/17/2021"
output:
  word_document: default
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
options(dplyr.summarise.inform = FALSE)
library(ggplot2)
library(knitr)
```

##question 1

Let $X_1,X_2,...X_n$ be independent, identically distributed random variables with mean $\mu$ and variance $\sigma^2$

$$\bar X=\frac{1}{n}\sum_{i=1}^nX_i$$

$$(X_i - \bar{X})^2 =  X_i^2 - 2X_i\bar{X}+\bar{X}^2$$
$$E\left[\sum_{i=1}^n\left(X_i-\bar X\right)^2\right]= E\left[\sum_{i=1}^n(X_i^2 - 2X_i\bar{X}+\bar{X}^2)\right]$$
$$E\left[\sum_{i=1}^n\left(X_i-\bar X\right)^2\right]= E\left[\sum_{i=1}^nX_i^2 - \sum_{i=1}^n2X_i\bar{X}+\sum_{i=1}^n\bar{X}^2\right]$$
$$E\left[\sum_{i=1}^n\left(X_i-\bar X\right)^2\right]=E\left[\sum_{i=1}^nX_i^2\right]-2E\left[\sum_{i=1}^n\bar XX_i\right]+E\left[\sum_{i=1}^n\bar X^2\right]$$
##Question 2
$$Var[X_i]=E[X_i^2]-E[X_i]^2$$
$$Var[X_i]=\sigma^2$$ 

$$E[X_i]=\mu$$
$$\sigma^2= E[X_i^2]-E[X_i]^2$$
$$\sigma^2= E[X_i^2]-\mu^2$$
$$ E[X_i^2]= \sigma^2+\mu^2$$
If $\mu=6$ and $\sigma^2=12$, then $ E [X_i^2]= 12+ 6^2$  so $ E [X_i^2]= 48$
```{r}
f2<-function(x){x^2*dgamma(x,shape=3,scale=2)}
integrate(f2,0,Inf)$value


```

## Question 3

$$E[X_i^2]=\sigma^2+\mu^2$$
$$E\left[\sum_{i=1}^nY_i\right]=\sum_{i=1}^nE[Y_i]$$
$$E\left[\sum_{i=1}^nX_i^2\right]=E[X_i^2]=\sigma^2+\mu^2$$

## Question 4

$$\bar X=\frac{1}{n}\sum_{i=1}^nX_i=\mu$$
$$Var[X_i]=\frac{\sigma^2}{n}$$
$$E\left[\sum_{i=1}^nY_i\right]=\sum_{i=1}^nE[Y_i]$$
$$E\left[\sum_{i=1}^n\bar X^2\right]= \sum_{i=1}^nE[\bar X^2]$$
$$E\left[\sum_{i=1}^n\bar X^2\right]=n[\mu^2+ \frac{\sigma^2}{n}]$$
$$E\left[\sum_{i=1}^n\bar X^2\right]=n\mu^2+ \sigma^2$$

## Question 5
$$E\left[\sum_{i=1}^n\bar XX_i\right]=E\left[\bar X\sum_{i=1}^nX_i\right]$$
$$\sum_{i=1}^nX_i=n\bar{X}$$

$$E\left[\sum_{i=1}^n\bar XX_i\right]=E\left[n\bar X^2\right]$$

## Question 6

$$E\left[\sum_{i=1}^nX_i^2\right]=n\left(\sigma^2+\mu^2\right)$$
$$E\left[\sum_{i=1}^n\bar X^2\right]=n\left(\frac{\sigma^2}{n}+\mu^2\right)$$
$$E\left[\sum_{i=1}^n\bar XX_i\right]=E\left[n\bar X^2\right]=n\left(\frac{\sigma^2}{n}+\mu^2\right)$$
$$E\left[\sum_{i=1}^nX_i^2\right]-2E\left[\sum_{i=1}^n\bar XX_i\right]+E\left[\sum_{i=1}^n\bar X^2\right]=n\left(\sigma^2+\mu^2\right) -2n\left(\frac{\sigma^2}{n}+\mu^2\right) + n\left(\frac{\sigma^2}{n}+\mu^2\right)$$
$$E\left[\sum_{i=1}^nX_i^2\right]-2E\left[\sum_{i=1}^n\bar XX_i\right]+E\left[\sum_{i=1}^n\bar X^2\right]=n\sigma^2+n\mu^2 -2\sigma^2 -2n\mu^2 + \sigma^2+n\mu^2$$
$$E\left[\sum_{i=1}^nX_i^2\right]-2E\left[\sum_{i=1}^n\bar XX_i\right]+E\left[\sum_{i=1}^n\bar X^2\right]=(n-1)\sigma^2$$

## Question 7

$$E\left[\frac{1}{n-1}\sum_{i=1}^n\left(X_i-\bar X\right)^2\right]=\frac{1}{n-1}E\left[\sum_{i=1}^n\left(X_i-\bar X\right)^2\right]$$

$$\frac{1}{n-1}E\left[\sum_{i=1}^n\left(X_i-\bar X\right)^2\right]=\frac{1}{n-1} \left[ E\left[\sum_{i=1}^nX_i^2\right]-2E\left[\sum_{i=1}^n\bar XX_i\right]+E\left[\sum_{i=1}^n\bar X^2\right]\right]$$

$$\frac{1}{n-1} \left[ E\left[\sum_{i=1}^nX_i^2\right]-2E\left[\sum_{i=1}^n\bar XX_i\right]+E\left[\sum_{i=1}^n\bar X^2\right]\right]=\frac{1}{n-1}(n-1)\sigma^2$$

$$E\left[\frac{1}{n-1}\sum_{i=1}^n\left(X_i-\bar X\right)^2\right]=\sigma^2$$


## Question 8
Part 1
$$f_x(0)=\frac{1}{2}$$
$$f_x(1)=\frac{1}{2}$$
$$f_y(2)=\frac{1}{3}$$
$$f_y(3)=\frac{2}{3}$$

$f_{x,y}(0,2)=f_x(0)*f_y(2)=\frac{1}{2}*\frac{1}{3}=\frac{1}{6}$ for a

$f_{x,y}(0,3)=f_x(0)*f_y(3)=\frac{1}{2}*\frac{2}{3}=\frac{1}{3}$ for b and c

$f_{x,y}(1,2)=f_x(1)*f_y(2)=\frac{1}{2}*\frac{1}{3}=\frac{1}{6}$ for d

$f_{x,y}(1,3)=f_x(1)*f_y(3)=\frac{1}{2}*\frac{2}{3}=\frac{1}{3}$ for e and f


 The random variables are independent because the probability of both events is the same as the probability of a single event occurring. 

Part 2
$f_x(0)=\frac{1}{2}$

$f_x(1)=\frac{1}{2}$

$f_y(2)=\frac{1}{3}$

$f_y(3)=\frac{2}{3}$

$f_{x,y}(0,2)=f_x(0)*f_y(2)=\frac{1}{2}*\frac{1}{3}=\frac{1}{6}$ for a and d

$f_{x,y}(0,3)=f_x(0)*f_y(3)=\frac{1}{2}*\frac{2}{3}=\frac{1}{3}$ for e

$f_{x,y}(1,3)=f_x(1)*f_y(3)=\frac{1}{2}*\frac{2}{3}=\frac{1}{3}$ for b and c and f


The random variables are not independent because the probability of both events is not the same as the probability of a single event occurring.

##Question 9
weighted average $xa+(1-x)b$ with $x\in[0,1]$ minimizes the variance of $xa+(1-x)b$

$var[xa+(1-x)b] =var[xa]+ var[(1-x)b]$

$=x var[a]+ (1-x)^2 var[b]$

$=x^2v +(1-x)^2 w$

$=(w+v)x^2-2wx+w$

$a=(w+v)$

$b=-2w$

$x=-\frac{b}{2a}$

$x=\frac{2w}{2(w+v)}$

$x=\frac{w}{(w+v)}$


